Comparative Analysis of Single--Neuron versus Dual--Neuron Output Layers in Binary Classification Neural Networks

### 2025-05-13 paper/model_architectures.tex
architecture diagrams added

Our custom Small CNN architecture represents a lightweight convolutional neural network designed specifically for this research. Figure~\ref{fig:small_cnn_arch} illustrates its structure with both single-neuron and dual-neuron output configurations.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{figures/small_cnn_1neuron_architecture.png}
\caption{Small CNN architecture with single-neuron output layer. The network consists of three convolutional blocks followed by fully connected layers and a single sigmoid output neuron.}
\label{fig:small_cnn_arch}
\end{figure}

@@

For our second architecture, we employed the ResNet50 model with transfer learning. Figure~\ref{fig:resnet50_arch} shows the architecture with its characteristic residual connections.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{figures/resnet50_1neuron_architecture.png}
\caption{ResNet50 architecture with single-neuron output layer. The model utilizes pre-trained weights from ImageNet, with only the final residual block and classifier layers fine-tuned for our binary classification tasks.}
\label{fig:resnet50_arch}
\end{figure}

@@

Our third architecture is the Vision Transformer (ViT), which represents the attention-based paradigm in computer vision. Figure~\ref{fig:vit_arch} illustrates the ViT architecture with its distinctive transformer encoder blocks.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{figures/vit_1neuron_architecture.png}
\caption{Vision Transformer (ViT) architecture with single-neuron output layer. The model divides the input image into patches, processes them through transformer encoder blocks, and uses the class token representation for classification.}
\label{fig:vit_arch}
\end{figure}

@@

For each architecture, we implemented two output layer configurations:

>>>>

Our custom Small CNN architecture represents a lightweight convolutional neural network designed specifically for this research. Figures~\ref{fig:small_cnn_arch_single} and \ref{fig:small_cnn_arch_dual} illustrate its structure with single-neuron and dual-neuron output configurations, respectively.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{figures/small_cnn_1neuron_architecture.png}
\caption{Small CNN architecture with single-neuron output layer. The network consists of three convolutional blocks followed by fully connected layers and a single sigmoid output neuron.}
\label{fig:small_cnn_arch_single}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{figures/small_cnn_2neuron_architecture.png}
\caption{Small CNN architecture with dual-neuron output layer. The network maintains the same structure as the single-neuron variant but uses two output neurons with softmax activation.}
\label{fig:small_cnn_arch_dual}
\end{figure}

@@

For our second architecture, we employed the ResNet50 model with transfer learning. Figures~\ref{fig:resnet50_arch_single} and \ref{fig:resnet50_arch_dual} show the architecture with its characteristic residual connections for both output configurations.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{figures/resnet50_1neuron_architecture.png}
\caption{ResNet50 architecture with single-neuron output layer. The model utilizes pre-trained weights from ImageNet, with only the final residual block and classifier layers fine-tuned for our binary classification tasks.}
\label{fig:resnet50_arch_single}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{figures/resnet50_2neuron_architecture.png}
\caption{ResNet50 architecture with dual-neuron output layer. The model maintains the same structure as the single-neuron variant but uses two output neurons with softmax activation.}
\label{fig:resnet50_arch_dual}
\end{figure}

@@

Our third architecture is the Vision Transformer (ViT), which represents the attention-based paradigm in computer vision. Figures~\ref{fig:vit_arch_single} and \ref{fig:vit_arch_dual} illustrate the ViT architecture with its distinctive transformer encoder blocks for both output configurations.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{figures/vit_1neuron_architecture.png}
\caption{Vision Transformer (ViT) architecture with single-neuron output layer. The model divides the input image into patches, processes them through transformer encoder blocks, and uses the class token representation for classification.}
\label{fig:vit_arch_single}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{figures/vit_2neuron_architecture.png}
\caption{Vision Transformer (ViT) architecture with dual-neuron output layer. The model maintains the same structure as the single-neuron variant but uses two output neurons with softmax activation.}
\label{fig:vit_arch_dual}
\end{figure}

@@

As shown in Figures \ref{fig:small_cnn_arch_single}--\ref{fig:vit_arch_dual}, for each architecture, we implemented two output layer configurations:
